---
title: "Regional Partner Annual Survey 2024-25"
author: "NWRI D Team"
output-dir: folder1
format:
  html:
    toc: true
    css: styles.css
editor: visual
---

```{r, echo=FALSE, include=FALSE, warning=FALSE, cache=FALSE}
require(tidyverse)
regional_Data<-readRDS("data/Regional_Partner_Survey2425.rds")
responders<-regional_Data %>%
  filter(!is.na(Q5_NPS_GROUP))  %>%
  nrow()
#calculate NPS
NPS_reg<-regional_Data %>%
  filter(!is.na(Q5_NPS_GROUP)) %>%
  count(Q5_NPS_GROUP) %>%
  ungroup() %>%
  mutate(perc = n/sum(n)) %>% 
  filter(Q5_NPS_GROUP == "Promoter") %>%
  mutate(NPS = round(perc * 100))

#calcualte customer satisfaction
cust_satis<-regional_Data %>%
  filter(!is.na(Q5_NPS_GROUP)) %>%
  mutate(satisfied  = ifelse(Q5 %in% c(8,9,10), "Satisfied", "Not")) %>%
  count(satisfied) %>%
  ungroup() %>%
  mutate(cust_perc = round(100*n/sum(n))) %>%
  filter(satisfied == "Satisfied")

#no detractors so the NPS is just the percentage of promotors


```
# Overview

We had a total of `{r} responders` responses to the regional partner annual survey. The overall NPS was `{r} NPS_reg$NPS ` with a customer satisfaction rate of `{r} cust_satis$cust_perc`%.

# Customer satisfaction rate info
We considered responses of 8-10 for the NPS questions as satisfied customers and calculated the percentage of those selecting 8-10 as the customer satisfaction percentage.

# Survey content
Below is a table showing each partner, their NPS score, and their fill in the blank answers to questions regarding why they chose their score and what sort of improvements they would like:
```{r, echo=FALSE, include=TRUE, warning=FALSE, cache=FALSE, message=FALSE}
require(reactable)
require(reactablefmtr)

paste(attr(regional_Data$Q7, "label"))

regional_table <- regional_Data %>%
  select(-Q5_NPS_GROUP) %>%
  select(Q2, Q5, Q6, Q7) %>%
  rename(
    Partner = Q2,
    `NPS Score` = Q5
  ) %>%
  rename_with(~ attr(regional_Data$Q6, "label"), .cols = Q6) %>%
  rename_with( ~ attr(regional_Data$Q7, "label"), .cols = Q7)

reactable(regional_table,
          filterable = TRUE,
          sortable = TRUE,
          wrap = TRUE)

```
