---
title: "Regional Partner Quarterly Report 2024-25"
author: "NWRI D Team"
output-dir: folder1
format:
  html:
    toc: true
    toc-location: left
    toc-title: "Contents"
    page-layout: full
    toc-depth: 2
    toc-floating: true
editor: visual
---

# DATA OVERVIEW

```{r,include = FALSE}
#load libraries
require(tidyverse)
require(sf)
require(kableExtra)
```

```{r, message=FALSE,include = FALSE}

dataq4<-readRDS("data/2024_2025/survey_data_q4_2425.rds")
dataq3<-readRDS("data/2024_2025/survey_data_q3_2425.rds")
dataq2<-readRDS("data/2024_2025/survey_data_q2_2425.rds")
dataq1<-readRDS("data/2024_2025/survey_data_q1_2425.rds")

regional_dataq4<-dataq4$regional
regional_dataq3<-dataq3$regional
regional_dataq2<-dataq2$regional
regional_dataq1<-dataq1$regional

rm(dataq4)
rm(dataq3)
rm(dataq2)
rm(dataq1)

#get some basic numbers up front for Q4
total_responders<-nrow(regional_dataq4)
total_partners<- regional_dataq4 %>%
  group_by(Q1) %>%
  count() 
total_NPS<-regional_dataq4 %>%
  group_by(Q20) %>%
  count() %>%
  rename(NPS_question = Q20,
         Responders = n) %>%
  ungroup() %>%
  mutate(Response_percent = 100 - round(100*(Responders/sum(Responders)),1))

NA_peeps<-total_NPS %>%
  filter(is.na(NPS_question))

#this is the percent response rate to the NPS question
NA_peeps$Response_percent
#calculate the total number of people responding to the NPS question 
responded_nps_total<-total_responders-NA_peeps$Responders

regional_data_timeq4<-regional_dataq4 %>%
  mutate(Quarter = "Quarter 4")
regional_data_timeq3<-regional_dataq3 %>%
  mutate(Quarter = "Quarter 3")
regional_data_timeq2<-regional_dataq2 %>%
  mutate(Quarter = "Quarter 2")
regional_data_timeq1<-regional_dataq1 %>%
  mutate(Quarter = "Quarter 1")
```

```{r,include = FALSE}
#get net promoter score for all


get_col_nps<-function(x){
  #first extract labels for the columns
column_labels <- sapply(x, function(col) as.character(attr(col, "label")))
questions<-tibble(column = colnames(x), 
           labels = column_labels) %>%
  mutate(labels = gsub("\\s+", " ", labels))
#now find which column has the NPS question
NPS_col<-which(str_detect(questions$labels, 
                 "likely are you to recommend New Worlds Reading to others?"))
if(length(NPS_col)>1){
  multicols<-questions[NPS_col,]
  singcol<-multicols %>%
    mutate(str_count = str_count(column)) %>%
    filter(str_count == min(str_count))
  NPS_col<-singcol$column
  }else{
  print("Only one type of question for NPS")
    NPS_col<-colnames(x)[NPS_col]
  }
return(NPS_col)}

NPS_colQ4<-get_col_nps(regional_dataq4)
NPS_colQ3<-get_col_nps(regional_dataq3)
NPS_colQ2<-get_col_nps(regional_dataq2)
NPS_colQ1<-get_col_nps(regional_dataq1)

calc_nps_total<-function(x,y){

  NPS_labels<-x %>%
  dplyr::select(all_of(y)) %>%
  mutate(across(everything(), ~ as.numeric(.))) %>%
  mutate( NPS_CAT = ifelse(.data[[y]] >= 9,
                           "Promoter",
                           ifelse(.data[[y]] <= 6,"Detractor",
                                  "Passive"))) %>%
  cbind(.,x$Q1)
Total_NPS = NPS_labels%>%
  drop_na() %>%
  group_by(NPS_CAT) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  mutate(prop = 100*(n/sum(n))) %>%
  pivot_wider(id_cols =!n,names_from = "NPS_CAT", values_from = "prop") %>%
  mutate(NPS = round(Promoter - Detractor,2)) 
return(Total_NPS)
}

Total_NPSQ4<-calc_nps_total(regional_dataq3, y = NPS_colQ4)
Total_NPSQ3<-calc_nps_total(regional_dataq3, y = NPS_colQ3)
Total_NPSQ2<-calc_nps_total(regional_dataq2, y = NPS_colQ2)
Total_NPSQ1<-calc_nps_total(regional_dataq1, y = NPS_colQ1)


Q4_totals<-regional_data_timeq4 %>%
  summarise(total_n = n(),
            total_answer = sum(!is.na(Q20)),
            total_answer_perc = round(100*(total_answer/total_n),1),
            NPS = Total_NPSQ4$NPS) %>%
  dplyr::select(-total_answer)


Q3_totals<-regional_data_timeq3 %>%
  summarise(total_n = n(),
            total_answer = sum(!is.na(Q20)),
            total_answer_perc = round(100*(total_answer/total_n),1),
            NPS = Total_NPSQ3$NPS) %>%
  dplyr::select(-total_answer)

Q2_totals<-regional_data_timeq2 %>%
  summarise(total_n = n(),
            total_answer = sum(!is.na(Q20)),
            total_answer_perc = round(100*(total_answer/total_n),1),
            NPS = Total_NPSQ2$NPS) %>%
  dplyr::select(-total_answer)

Q1_totals<-regional_dataq1 %>%
   summarise(total_n = n(),
            total_answer = sum(!is.na(Q20)),
            total_answer_perc = round(100*(total_answer/total_n),1),
            NPS = Total_NPSQ1$NPS) %>%
  dplyr::select(-total_answer)


Q4_totals<-Q4_totals %>%
  mutate(Quarter = 4)

Q3_totals<-Q3_totals %>%
  mutate(Quarter = 3)

Q2_totals<-Q2_totals %>%
  mutate(Quarter = 2)

Q1_totals<-Q1_totals %>%
  mutate(Quarter = 1)

temporal_nps<-rbind(Q1_totals,Q2_totals,Q3_totals, Q4_totals) 

#difference in NPS 
diff_nps <- abs(temporal_nps$NPS[temporal_nps$Quarter == 4]-temporal_nps$NPS[temporal_nps$Quarter == 3])

perc_diff<-round(100*(diff_nps/temporal_nps$NPS[temporal_nps$Quarter == 4]),1)

#change in respondents
total_change = Q4_totals$total_n - Q3_totals$total_n

total_NPS<-regional_data_timeq4 %>%
  drop_na(Q20)%>%
  group_by(Q20) %>%
  count() %>%
  rename(NPS_question = Q20,
         Responders = n) %>%
  ungroup() %>%
  mutate(Response_percent = round(100*(Responders/sum(Responders)),1))

```

```{r,include=FALSE}
#calculating percent of NPS question answered 
regional_part_df<-regional_data_timeq4 %>%  drop_na(Q1) %>%
  group_by(Q1) %>% 
  summarise(n = n(),
            answerednps = sum(!is.na(Q20))) %>%
  ungroup() %>%
  mutate(Percent =round(100*(answerednps/n),1)) %>%
  dplyr::select(-answerednps) %>%
  arrange(desc(Percent))


#calc NPS by regional partner:

reg_npsQ4<- regional_data_timeq4 %>%
  dplyr::select(all_of(NPS_colQ4)) %>%
  mutate(across(everything(), ~ as.numeric(.))) %>%
  mutate( NPS_CAT = ifelse(.data[[NPS_colQ4]] >= 9,
                           "Promoter",
                           ifelse(.data[[NPS_colQ4]] <= 6,"Detractor",
                                  "Passive"))) %>%
  cbind(.,regional_data_timeq4$Q1) %>%
  rename(partner = `regional_data_timeq4$Q1`) %>%
    drop_na() %>%
  group_by(partner, NPS_CAT) %>%
  summarise(n = n()) %>%
  group_by(partner) %>%
  mutate(prop = 100*(n/sum(n))) %>%
    pivot_wider(id_cols ="partner",names_from = "NPS_CAT", 
                values_from = "prop",
                values_fill = 0) %>%
  mutate(NPS = round(Promoter - Detractor,2)) %>%
  dplyr::select(partner, NPS)


regional_part_df_tableQ4<-left_join(regional_part_df,reg_npsQ4, by = c("Q1" = "partner")) %>%
  arrange(desc(NPS))

#maximum nps partners
maxnps_reg<-regional_part_df_tableQ4 %>% filter(NPS == max(regional_part_df_tableQ4$NPS, na.rm = T)) %>%
  dplyr::select(Q1) 

minnps_reg<-regional_part_df_tableQ4 %>% filter(NPS == min(regional_part_df_tableQ4$NPS,na.rm = T)) %>%
  dplyr::select(Q1) 


```

```{r,include=FALSE}
#calculating percent of questions answered 
percents<- regional_data_timeq4 %>%
  select(starts_with("Q")) %>%
  mutate(percent_na = rowMeans(is.na(.)) * 100) %>%
  ungroup()
percentabove50=100*round(length(which(percents$percent_na>50))/nrow(percents),1)

#month with the most responses
max_response<-regional_data_timeq4 %>%
   mutate(week = paste(year(RecordedDate), sprintf("%02d", week(RecordedDate)), sep = "-"),
         month = floor_date(RecordedDate, "month"),
         month_label = factor(format(month, "%b"), 
                              levels = month.abb),
         year_month = format(month, "%Y-%m"))%>%
  group_by(month_label) %>%
  count() 

max_response_month <- as.character(max_response$month_label[max_response$n == max(max_response$n)])
max_response_number <- max(max_response$n)

```

## Quarter 4 Overview:

-   Total survey respondents: `{r} total_responders`
    -   `{r} responded_nps_total` (`{r} NA_peeps$Response_percent`%) respondents answered the NPS question.
    -   `{r} length(which(percents$percent_na>50))` (`{r} percentabove50`%) respondents answered 50% or more of the survey questions.
-   Overall NPS: `{r} Total_NPSQ3$NPS`
    -   Highest NPS: `{r} max(regional_part_df_tableQ4$NPS,na.rm =T)` By: `{r} maxnps_reg$Q1`
    -   Lowest NPS: `{r} min(regional_part_df_tableQ4$NPS,na.rm =T)` By: `{r} minnps_reg$Q1`
-   The month with the most responses was `{r} max_response_month` with `{r} max_response_number` responses.

## Quarter 4 compared to the previous Quarter:

-   There were `{r} abs(total_change)` `{r} if (total_change >= 0) "more" else "less"` respondents.
-   The NPS score saw a `{r} if (diff_nps >= 0) "increase" else "decrease"` of `{r} diff_nps` (`{r} perc_diff`%) (Table 1).

```{r, echo=FALSE, include=TRUE, warning=FALSE, cache=FALSE}
#| out-width: 65%
#| #| label: fig_histogram
#| fig-cap: "Number of survey responses by month"

response_time <- rbind(regional_data_timeq2, regional_data_timeq1, regional_data_timeq3, regional_data_timeq4) %>%
  mutate(week = paste(year(RecordedDate), sprintf("%02d", week(RecordedDate)), sep = "-"),
         month = floor_date(RecordedDate, "month"),
         month_label = factor(format(month, "%b"), 
                              levels = month.abb),
         year_month = format(month, "%Y-%m")) %>%
  group_by(Quarter, year_month) %>%
  summarise(n = n(), .groups = "drop") %>%
  ggplot() +
  geom_col(aes(x = year_month, y = n, fill = Quarter)) +
  theme_light() +
  labs(y = "No. of Survey Responses") +
  theme(legend.title = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 40, vjust = 0.5)) +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_viridis_d(option = "E")
response_time


```

```{r, echo=FALSE, include=TRUE, warning=FALSE, cache=FALSE}

kableExtra::kable(temporal_nps,"pipe",
                  col.names = c("Total Survey Respondents",
                                "Response rate of NPS question(%)",
                                "Net Promoter Score", "Quarter"),
                  align = "c",
                  caption = "Numbers on: survey response, NPS question not answered, and net promoter score across quarters") %>%
  kableExtra::kable_styling(latex_options = "striped")

```

Additionally across the `{r} length(na.omit(total_partners$Q1))` regional partners, the distribution of these respondents to the NPS question and the NPS for each partner in Q3 are as follows:

```{r, echo=FALSE, include=TRUE, warning=FALSE, cache=FALSE}
kableExtra::kable(regional_part_df_tableQ4,"pipe",
                  col.names = c("Regional Partner", 
                                "Survey Takers", 
                                "Response rate to NPS question (%)",
                                "NPS"),
                  align = "c",
                  caption = "Table showing NPS across regional partners")%>%
  kableExtra::kable_styling(latex_options = "striped") %>%
kableExtra::column_spec(3, width = "3cm")
  

```

# Q4 Data Dive

## Details on NPS across regional partners

The frequency of the possible answers selected, ranging from 0 to 10, for the NPS question across the six largest regional partners (based on respondent numbers) are shown in the figure below for Q2.

```{r, echo=FALSE, include=TRUE, warning=FALSE, cache=FALSE}
#| out-width: 100%
#| #| label: fig_histogram
#| fig-cap: "Distribution of scores (0-10) to the net promotor question from the survey across the top 6 regional partners in terms of respondent numbers. Orange = 0-6, blue = 6-9, green = 9-10"

regional_part_df4<-regional_dataq4 %>%  drop_na(Q1) %>%
  group_by(Q1) %>% count() %>%
  ungroup() %>%
  mutate(Percent =100*round(n/sum(n),1)) %>%
  arrange(desc(n)) %>%
  slice(1:6)

plotdata<-regional_dataq4 %>%
  drop_na(Q1) %>%
  filter(Q1 %in% as.character(regional_part_df4$Q1)) 
ggplot(data = plotdata) +
  scale_x_continuous(breaks =scales::breaks_pretty()) +
  labs(y = "Responder count", x = "NPS question scores") +
  annotate("rect",xmin=0, xmax=6, ymin=0, ymax=Inf, fill = "#E69F00",
              alpha = 0.5 )+
    annotate("rect",xmin=6, xmax=9, ymin=0, ymax=Inf, fill = "#56B4E9",
              alpha = 0.5 )+
    annotate("rect",xmin=9, xmax=10, ymin=0, ymax=Inf, fill = "#009E73",
              alpha = 0.5 )+
    geom_histogram(data = plotdata, aes(x = as.integer(Q20)), binwidth = 1) +
  facet_wrap(~Q1, scales = "free_y",
             labeller =label_wrap_gen(width=20)) +
  theme_bw() 
```

```{r,include=FALSE, warning=FALSE}
#let's look and see how many people do not have an enrolled child
regional_data_eleigibility<-regional_dataq4 %>%
  group_by(Q6) %>%
  count() %>%
  ungroup() %>% mutate(percent = round(100*(n/sum(n)),1))
```

While most people selected 10 for the NPS question certain regional partners had higher frequencies of passive scores (6-9, bars in the blue region of Figure 2). Detractors (0-6) were sparse but still appeared in almost all regional partner events.

## Details on how likely respondents would enroll their child in NWRI

Out of all responses, `{r} regional_data_eleigibility$percent[regional_data_eleigibility$Q6 == "No"& !is.na(regional_data_eleigibility$Q6)]`% (`{r} regional_data_eleigibility$n[regional_data_eleigibility$Q6 == "No"& !is.na(regional_data_eleigibility$Q6)]` respondents) did not have a child enrolled in NWRI.

```{r,include=FALSE}
#responders who do not have children in the NWRI
responder_no_nwri<-regional_dataq3 %>%
  filter(Q6 == "No") %>%
  group_by(`Q6-2`) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  mutate(percent = round(100*(n/sum(n)),1))

kableExtra::kable(responder_no_nwri,"pipe",
                  col.names = c("Would you enroll?", "No.", "%"),
                  align = "c",
                  caption = "Responses from people with no child in NWRI when asked if they would they enroll their child in NWRI?") %>%
   kableExtra::kable_styling(latex_options = "striped")
```

When these `{r} regional_data_eleigibility$n[regional_data_eleigibility$Q6 == "No"& !is.na(regional_data_eleigibility$Q6)]` respondents were asked if they would consider enrolling their child the responses were as follows:

Note! We auto-detected their responses by matching key words to categorize responses into several groups. Because this approach relies on key words matching not all the responses could be properly assigned a category.

```{r, echo=FALSE, include=TRUE, warning=FALSE, cache=FALSE}
reasons_why<-regional_dataq4 %>%
  filter(Q6 == "No" & `Q6-2` == "No") %>%
  dplyr::select(Q7) %>%
  drop_na()

wtf<-reasons_why %>%
  mutate(Q7 = str_to_lower(Q7),
         Q7 = str_remove_all(Q7, "[[:punct:]]")) %>%
  mutate(recat = case_when(
    str_detect(Q7, paste(
      "doesnt qualify", "didnt qualify", "dont qualify", 
      "ineligible", "above reading level", "not qualify", 
      "aboved grade level","not old enough","not eligible",
      "not going to qualify", "scores were too high",
      sep = "|")) ~ "Doesn't qualify",
    
    str_detect(Q7, paste(
      "aged out", "age out", "too old", "to old", 
       "in high school", "in middle school",
      sep = "|")) ~ "Aged out",
    
     str_detect(Q7, paste(
      "dont have children", "no child", "no children", "childless", 
      sep = "|")) ~ "No children"
    
  ))

reasons_whycat<-reasons_why %>%
  mutate(Q7 = str_to_lower(Q7),
         Q7 = str_remove_all(Q7, "[[:punct:]]")) %>%
  mutate(Reason = case_when(
    str_detect(Q7, paste(
      "doesnt qualify", "didnt qualify", "dont qualify", 
      "ineligible", "above reading level", "not qualify", 
      "aboved grade level","not old enough","not eligible",
      "not going to qualify","scores were too high",
      "doesnt meet the eligibility", "reads above grade level",
      "reading above their grade", "arent eligible", 
      "requirement not met",
      
      sep = "|")) ~ "Doesn't qualify or not eligible",
    
    str_detect(Q7, paste(
      "aged out", "age out", "too old", "to old", 
       "in high school", "in middle school", "6th grade", "5th grade",
      sep = "|")) ~ "Explicitly mentioned aged out",
    
     str_detect(Q7, paste(
      "dont have children", "no child", "no children", "childless", 
      "dont have small kids", "dont have kids",
      sep = "|")) ~ "Explicitly mentioned no children"
    
  )) %>%
  group_by(Reason) %>%
  count() %>%
  mutate(Reason = ifelse(is.na(Reason), "Unmatched responses", Reason))

kableExtra::kable(reasons_whycat,"pipe",
                  col.names = c("Reason", "No. of responses"),
                  align = "c",
                  caption = "Why will you not enroll your child in NWRI?") %>%
  kableExtra::kable_styling(latex_options = "striped")
  
```

Across regional partners most respondents without an enrolled NWRI child are likely to recommend NWRI. However certain regional partners have a proportion of their respondents saying they were: unsure, preferred not to answer, or they were not likely to recommend NWRI. Figure 3 shows which partners may have higher proportions of these respondents (color in purple).

```{r include=FALSE}
regional_numbers<-regional_dataq4 %>%
  filter(Q6 == "No") %>%
  rename(willapply = `Q6-2`) %>%
  mutate(willapply = as.character(willapply)) %>%
  mutate(willapply = ifelse(is.na(willapply), "No Answer",willapply)) %>%
  count(Q1) %>%
  rename(respondents_n = n)
```

```{r, echo=FALSE, include=TRUE, warning=FALSE, cache=FALSE}
#| out-width: 100%
#| label: fig_comp_plot
#| fig-cap: "Response when respondents with no child enrolled in NWRI would consider enrolling their child. Total numbers reflect the total amount of respondents."

part_plot<-regional_dataq4 %>%
  filter(Q6 == "No") %>%
  rename(willapply = `Q6-2`) %>%
  mutate(willapply = as.character(willapply)) %>%
  mutate(willapply = ifelse(is.na(willapply), "No Answer",willapply)) %>%
  count(Q1,willapply) %>%
  group_by(Q1) %>% 
  mutate(perc = round(100*(n/sum(n)),1)) %>%
  left_join(., regional_numbers, by = "Q1")

ggplot(data =NULL) +
  geom_col(data =part_plot, aes(x = Q1, y = perc, fill = willapply)) +
  geom_text(data = regional_numbers, aes(x = Q1, y= 0,
                                         label = paste("Total:",respondents_n)),
            hjust = 0) +
  coord_flip() +
   theme_bw() +
  scale_fill_viridis_d(option = "D") +
  labs(y = "Response proportions (%)") +
   theme(axis.title.y =  element_blank(),
         legend.position = "right",
         plot.caption.position = "plot") +
    labs(fill = "Response") +
  geom_hline(aes(yintercept = 25), lty = 3, linewidth = 0.5, color = "grey") +
    geom_hline(aes(yintercept = 50), lty = 3, linewidth = 0.5, color = "grey") +
  geom_hline(aes(yintercept = 75), lty = 3, linewidth = 0.5, color = "grey") 

```
